\documentclass[a4paper,oneside,12pt]{book}

%----------------------------------------------------------------------------------------
%	README!
%   Welcome. It's worth having a read through this file
%   to set up the broad parameters, such as the name of
%   the degree, the school/department, the type of work
%   (dissertation/Final Year Project/report, etc. as well
%   as your own details.
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	COVER PAGE
%   The cover page is laid out in title/title.tex. You can choose a colour
%   or black and white logo
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------
%	THESIS INFORMATION
%   Put title, author name, supervisor name, degree, type of work, school, department in here
%   It will be used for the title page and for the embedded PDF information
%----------------------------------------------------------------------------------------

\newcommand{\thesistitle}{Development and Comparative Evaluation of a Fine-Tuned Multimodal LLM Against Base Model and ChatGPT-4} % Your thesis title, this is used in the title and abstract
\newcommand{\degree}{BAI (Computer Engineering)} % Replace with your degree name, this is used in the title page and abstract
\newcommand{\typeofthesis}{dissertation} % dissertation, Final Year Project, report, etc.
\newcommand{\authorname}{Swetang Krishna} % Your name, this is used in the title page and PDF stuff
%% Do not put your Student ID in the document, as TCD will not publish
%% documents that contain both your name and your Student ID.
\newcommand{\supervisor}{Prof. Vincent Wade} % replace with the name of your supervisor
%\newcommand{\cosupervisor}{Dr Alex Lee} % replace with the name of your co-supervisor if you have one
\newcommand{\keywords}{LLM, NLP, AI} % Replace with keywords for your thesis
\newcommand{\school}{\href{https://www.tcd.ie/Engineering/}{School of Engineering}} % Your school's name and URL, this is used in the title page
%Edited by HS for engineering

%% Comment out the next line if you don't want a department to appear
%\newcommand{\department}{\href{http://researchgroup.university.com}{Department Name}} % Your research group's name and URL, this is used in the title page


%% Language and font encodings
\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{float}

%% Bibliographical stuff
\usepackage[]{cite}
%% Document size
% include showframe as an option if you want to see the boxes
\usepackage[a4paper,top=2.54cm,bottom=2.54cm,left=2.54cm,right=2.54cm,headheight=16pt]{geometry}
\setlength{\marginparwidth}{2cm}
%% Useful packages
\usepackage{amsmath}
\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
\usepackage[pdftex]{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage[colorlinks=true, allcolors=black]{hyperref}
\usepackage{hyperxmp}
\usepackage{caption} % if no caption, no colon
\usepackage{sfmath} %use sans-serif in the maths sections too
\usepackage[parfill]{parskip}    % Begin paragraphs with an empty line rather than an indent
\usepackage{setspace} % to permit one-and-a-half or double spacing
\usepackage{enumerate} % fancy enumerations like (i) (ii) or (a) (b) and suchlike
\usepackage{booktabs} % To thicken table lines
\usepackage{fancyhdr}
\usepackage{xcolor} % to get TCD colour on headings
\numberwithin{equation}{chapter} %HS edit for (chapter.equation)
\pagestyle{plain} % Embrace simplicity!

\definecolor{tcd_blue}{RGB}{5, 105, 185}

%% It's personal taste but...
%% Uncomment the following block if you want your name and ID at the top of
%% (almost) every page.

%\pagestyle{fancy}
%\fancyhf{} % sets both header and footer to nothing
%\renewcommand{\headrulewidth}{0pt}
%\cfoot{\thepage}
%\ifdefined\authorid
%\chead{\it \authorname\ (\authorid)}
%\else
%\chead{\it \authorname}
%\fi
%% End of block

%% It is good practise to make your font sans-serif to improve the accessibility of your document.  Comment out the following line to disable it (but you really should not)
\renewcommand{\familydefault}{\sfdefault} %use the sans-serif font as default

%% If you insist on not using sans-serif (please don't), consider using Palatino instead of the LaTeX standard
%\usepackage{mathpazo} % Use the Palatino font by default if you prefer it to Computer Modern


%% Format Chapter headings appropriately
\usepackage{titlesec}
\titleformat{\chapter}[hang]{\normalfont\huge\bfseries\color{tcd_blue}}{\thechapter}{1cm}{}{}

\title{\thesistitle}
\author{\authorname}


\hypersetup{
   pdftitle=\thesistitle, % Set the PDF's title to your title
   pdfauthor=\authorname, % Set the PDF's author to your name
   pdfkeywords=\keywords, % Set the PDF's keywords to your keywords
   pdfsubject=\degree, % Set the PDF's keywords to your keywords
   pdfinfo={
     pdfsupervisor=\supervisor, % Set the PDF's supervisor to your supervisor
     %pdfcosupervisor=\cosupervisor, % Set the PDF's cosupervisor to your cosupervisor if using
   }
}


\frontmatter
\begin{document}
\input{title/title.tex}
\pagenumbering{roman}
\section*{\Huge\textcolor{tcd_blue}{Declaration}}
\vspace{1cm}
I hereby declare that this \typeofthesis\ is entirely my work and that it has not been submitted as an exercise for a degree at this or any other university.

\vspace{1cm}
I have read and understand the plagiarism provisions in the General Regulations of the University Calendar for the current year, found at \url{http://www.tcd.ie/calendar}.
\vspace{1cm}

I have completed the Online Tutorial on avoiding plagiarism `Ready Steady Write', located at \url{http://tcd-ie.libguides.com/plagiarism/ready-steady-write}.
\vspace{1cm}

I consent to the examiner retaining a copy of the thesis beyond the examining period, should they so wish (EU GDPR May 2018).
\vspace{1cm}

I agree that this thesis will not be publicly available, but will be available to TCD staff and students in the University’s open-access institutional repository on the Trinity domain only, subject to Irish Copyright Legislation and Trinity College Library conditions of use and acknowledgement.
\vspace{3cm}

Signed:~\rule{5cm}{0.3pt}\hfill Date:~\rule{5cm}{0.3pt}

\chapter*{Abstract}
This dissertation delves into the transformative potential of chatbots in the educational sector, highlighting their role in personalizing learning experiences and providing administrative support. With an emphasis on student engagement and personalized support, it examines the integration of AI-driven interfaces into educational settings. The research explores how chatbots can evolve learning environments into more interactive and student-focused platforms. Furthermore, this study extends into the domain of artificial intelligence, evaluating the effectiveness of question-answering (QA) systems and developing a fine-tuned multimodal Large Language Model (LLM) specifically for context-based QA in probability and statistics. The investigation centres on three pivotal areas: a base model pre-trained on QA tasks, a fine-tuned multimodal LLM, and its comparison with the advanced ChatGPT-4 model. The development of the fine-tuned LLM encompasses the enhancement of a base model through Retrieval-Augmented Generation (RAG) techniques and specialized training on a dataset curated from probability and statistics queries, aiming to rival or surpass the proficiency of larger models like ChatGPT-4, while also addressing the computational and financial challenges associated with such models.

This project's comprehensive approach to development and evaluation illuminates the significant impact of fine-tuning, multimodal capabilities, and RAG on improving the model's ability to process and understand complex, domain-specific questions. The findings offer an in-depth analysis of the efficacy, accuracy, and efficiency of the developed model in comparison to existing models, underscoring the nuanced advantages and limitations of these methodologies in enhancing QA performance. By providing a detailed framework for the development and assessment of fine-tuned multimodal LLMs, this dissertation contributes to the broader AI and education fields. It paves the way for future AI-driven educational technologies, emphasizing the role of specialized, efficient, and accessible LLMs in tackling intricate QA challenges within educational contexts.

\chapter*{Lay Abstract}
Imagine a classroom where every student gets immediate, personalized help on their studies, any time of the day - this is what chatbots in education can offer. This research project looks at how these smart, conversational computer programs, or chatbots, are changing the way we learn and manage educational tasks. They can tailor learning to each student's needs, answer questions instantly, and are always available, tackling common problems like keeping students interested and providing individual support.

Moreover, the study ventures further into the world of artificial intelligence by creating and testing a special kind of computer model designed to answer complex questions in a specific area: probability and statistics. This model is not just any model; it's built to be smarter and more efficient than some of the most advanced systems out there, like ChatGPT-4, but without needing as much computer power or money to run. This was achieved by starting with a basic model skilled in answering questions, enhancing it with advanced techniques, and training it on a special set of questions and answers.

The research aimed to find out how well this newly developed model can understand and answer tough questions compared to other big models. The results showed that by focusing on certain improvements, such as making the model work with different types of information and helping it retrieve relevant knowledge when needed, it could indeed handle complex, subject-specific questions very effectively.

This work doesn't just advance our understanding of artificial intelligence; it also opens up new possibilities for using smart technologies in education. It highlights how specialized, efficient computer models can be a powerful tool for overcoming challenges in learning and teaching, making high-quality education more accessible to everyone.



\newpage
\onehalfspacing\raggedright %\raggedright turns off justification and hyphenation

\section*{\Huge\textcolor{tcd_blue}{Acknowledgements}}
This project, culminating in May 2024, has been a journey of both challenge and discovery, none of which would have been possible without the unwavering support of several key individuals and groups. Foremost, I extend my deepest gratitude to Professor Vincent Wade, whose guidance and expertise were instrumental in navigating the complexities of this research. His patience, wisdom, and encouragement have left an indelible mark on my academic and personal growth.

To my family, whose belief in my capabilities never wavered, I owe a debt of gratitude. My parents, for their unconditional love and encouragement, and my sister, who has been my greatest cheerleader.

My heartfelt thanks go out to the friends I've been fortunate to make along this academic journey. Their camaraderie, support, and shared moments of stress and success have made this arduous journey a memorable one. 

To my peers in the Class of 2023 and 2024, who have navigated the challenges of these times with resilience and solidarity, I am proud to have shared this journey with you. The friendships forged and the collective spirit of our cohort have enriched this experience beyond measure.

Lastly, I acknowledge the broader academic and research community, whose contributions have paved the way for this project. Their collective knowledge and innovations have been invaluable to my research.

Thank you all for your part in this journey, for the lessons learned, and for the memories cherished. Your support has been a beacon of light in the pursuit of knowledge and excellence.

\tableofcontents
\listoffigures
\newpage

\mainmatter
% Maintaining separate .tex files for each chapter is good practice
\input{introduction/introduction.tex}
\input{background/background.tex}
\input{project/project.tex}
\input{implementation/implementation.tex}
\input{evaluation/evaluation.tex}
\input{conclusion/conclusion.tex}

% note that your supervisor may have a strong opinion on the style of referencing you use. Some background is available at https://www.overleaf.com/learn/latex/Bibtex_bibliography_styles
\bibliographystyle{IEEEtran} %Changed to IEEETran by HS
%\bibliographystyle{unsrt}

\bibliography{bibs/sample}
\appendix
\renewcommand{\thechapter}{A\arabic{chapter}}
\input{appendix/appendix.tex}


\end{document}